---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

This package is a collection of simple R functions that I've either written or discovered over the years that I've found useful and like to re-use. In some cases, these functions are just thin wrappers around other functions but the syntax is easier for me to remember.

This is still a work in progress.

To install, you need to have the devtools package on your system (if you don't have it, install that with install.packages("devtools") ) and then run

```{r, eval = FALSE}
devtools::install_github("smach/rmiscutils")
```

Some of the tasks that functions in this package accomplish:


Add a row to a dataframe with the format
```{r, eval = FALSE}
 add_row(mydf, col1 = 4, col2 = 7, col3="A")
```
`add_row`

Add a row to a dataframe with the format
```{r, eval = FALSE}
add_row_from_list(iris, list(5.1, 3.5, 1.4, 0.2, "setosa"))
```
`add_row_from_list`

Calculate percentages from selected data frame columns by row, and add new columns with that data. Designed for election results. `add_pct_cols_to_dataframe`

Calculate the statistical mode of a vector (which base R oddly does not do).
`mymode`

Capitalizes the first letter of each word in a string. (Note: base R 3.2.0 subsequently included the `tools::toTitleCase(`) function, but toinitialcap has the advantage of naming consistency with base R's `tolower` and `toupper`).
`toinitialcap` -- to go along with base R's tolower() and toupper(), 

Convert all NA values in a data frame to zeroes.
`na2zero`

Convert only NA values to zeroes in data frame columns that are numeric or integers. Uses for loops, so may be slow for large data frames.
`na2zero2`

Convert character strings containing numbers with commas into numbers.
`numbers_with_commas`

Convert a vector of character strings into an ordered factor, keeping the original order. Helpful if x-axis order is important when visualizing data. (Created before the forcats package).
`strings_to_ordered_factors`

Create a frequency table sorted by descending frequency. (Would probably use a version of `janitor::tabyl` today)
`table_sorted`

Find winning candidates in a spreadsheet or data frame of election results. `elec_find_winner`

Find winning (or other) streaks in a vector of data. This is a slightly modified function from [Exploring Baseball Data with R](https://baseballwithr.wordpress.com/2014/07/07/team-streaks-part-i-2/).
`streaks`

Read a text file into an R variable as a single character string.
`read_textfile`

Rename a single data frame column using format `renamecol(mydf, oldcolname, newcolname)`
`renamecol`

Remove columns from a data frame where _all_ values are NA. Useful when importing spreadsheets or CSV files where there are some empty columns.
`removecolsAllNA`

Unload a package from current working session. Because it can be difficult to remember that  `detach("package:mypackagename", unload=TRUE)` or `unloadNamespace("mypackagename")` are the opposites of `library("mypackagename")`.
`unlibrary`

Write an object containing character string data to a text file.
`write_textfile`

From a character string, return the value of an object with that name (there's also string2varname, which does the reverse)
`varname2string`

To see a complete list of available function and links to their help files, run

```{r, eval = FALSE}
help(package = "rmiscutils")
```

Or, to list some of the package's functions by function name:

*add_pct_cols_to_dataframe* -- designed for election results, will calculate percentages from selected columns by row and add new columns with that data. 

*add_row* -- add a row to a dataframe with the format
```{r, eval = FALSE}
 add_row(mydf, col1 = 4, col2 = 7, col3="A")
```
Thanks to Hadely Wickham for that function idea.

*add_row_from_list* -- add a row to a dataframe with the format
```{r, eval = FALSE}
add_row_from_list(iris, list(5.1, 3.5, 1.4, 0.2, "setosa"))
```
Thanks to Jennifer Bryan for that function idea.

*mymode* -- calculates the statistical mode of a vector, which base R oddly does not do.

*na2zero* -- turn all NA values in a data frame to zeroes.

*na2zero2* -- turns only NA values to zeroes in data frame columns that are numeric or integers. Uses for loops, so may be slow for large data frames.

*numbers_with_commas* -- easily turn character strings containing numbers with commas into numbers.

*read_textfile* -- read a text file into an R variable as a single character string.

*renamecol* -- rename a single column in a dataframe using format renamecol(mydf, oldcolname, newcolname)

*removecolsAllNA* -- remove columns from a data frame where _all_ values are NA. Useful when importing spreadsheets or CSV files where there are some empty columns.

*toinitialcap* -- to go along with base R's tolower() and toupper(), capitalizes the first letter of each word in a string. Note: base R 3.2.0 now has tools::toTitleCase() function.

*streaks* -- find winning (or other) streaks in a vector of data, slightly modified function from [Exploring Baseball Data with R](https://baseballwithr.wordpress.com/2014/07/07/team-streaks-part-i-2/).

*strings_to_ordered_factors* -- turn a vector of character strings into an ordered factor, keeping the original order. Helpful if x-axis order is important when visualizing data.

*table_sorted* -- get frequency table sorted by frequency descending.

*unlibrary* -- because it can be difficult to remember that  detach("package:mypackagename", unload=TRUE) or unloadNamespace("mypackagename") are the opposites of library("mypackagename").

*varname2string* -- takes string and returns value of an object with that name (there's also string2varname, which does the reverse)

*write_textfile* -- write an object containing character string data to a text file.

To see a complete list of available function and links to their help files, run

```{r, eval = FALSE}
help(package = "rmiscutils")
```
